Lens Modeling in \sw

### 2014-01-18 text/ms.tex
Correcting grammatical errors and adding some further clarification to the abstract.

\begin{abstract}
In Spacewarps, volunteers are invited to search CFHTLS sky survey for
lens candidates.  Here we report on a system that allows experienced
volunteers to model lens candidates.  A sample of 29 simulated lenses
were modeled, each multiple times.  The quality of the models was then
examined in different ways.  In particular: (i)~the image parities and
time arrivals were identified correctly in most cases, but not
always; (ii)~the mean convergence (equivalent to the enclosed mass)
was quite well constrained in the image region; (iii)~the performance
of experienced volunteers was comparable to professionals.
\end{abstract}

@@

In response to the lessons learned from automated lens searches, the
{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org} 
(Saha is a co-investigator) has been launched.  {\em Spacewarps\/} is
part of the {\em Zooniverse\/} family of citizen-science
projects,\footnote{\tt http://www.zooniverse.org} where members of the
public are invited to analyze different kinds of scientific data which
are too difficult for robots and too large for specialists.  In {\em
 Spacewarps\/} itself, since the launch in May 2013, $\sim100$ square
degrees (or a quarter of a percent of the sky) has been examined.  The
survey is divided into patches of $440\times 440$ pixels, each of
which is seen by ten volunteers.  Simulated lenses are mixed in with
the data, both to help train volunteers on what to look for, and to
estimate completeness and purity.  A list of candidates is being
processed, but intermediate results show some very good candidates, as
well re-discoveries of a few known lenses.

@@

The sims were produced using {\tt gravlens}
\citep{2001astro.ph..2341K,2001astro.ph..2340K}. They were of three
kinds, as follows.

\begin{enumerate}
  \item Imitating lensed quasars: having a singular elliptical
    isothermal lens (SIE) plus constant external shear, and a circular
    Gaussian source.
  \item Emulating lensed galaxies: similar to the above, but with an
    elliptical de Vaucouleurs source.
  \item Resembling cluster lenses: having a source as above, but a
    more complicated lens, with one dominant elliptical SIE and
    one or more perturbing elliptical SIEs, plus a circular NFW
    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
    the intergalactic mass.
\end{enumerate}

@@

To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate \ERg.
To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
These results can be seen in \figref{ER_all_models}.
This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.

@@

\begin{figure}
  \centering
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007020_kappa_encl}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007021_kappa_encl}} \\
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007024_kappa_encl}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007025_kappa_encl}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007022_kappa_encl}}
  \caption{\kenc for models of \asw{0h2m}: 7020 (top left), 7021 (top right), 7024 (mid left), 7025 (mid right) by volunteers and a correct model 7022 (bottom) by an expert.}
  \label{fig:kapenc_compare_faulty}
\end{figure}

@@

E05 and E06 happen mostly in very symmetric images, that are hard to come up with a unique valid solution.
This errors change the orientation of the mass distribution \kap[x,y],
but \kenc and thus the \ERf is not influenced that much, and thus the final result is still a valid model.
This can be seen by comparing the results for \asw{0h2m}, shown in \figref{kapenc_compare_faulty}: 
The correct model 7022 ($\ERm=10.76px$), which was done by an expert, and
7020 ($\ERm=10.72px$),
7024 ($\ERm=10.80px$),
7025 ($\ERm=11.16px$),
7021 ($\ERm=11.04px$).
The first three are of type E06, while the last is of type E05.
This can be easily corrected, if further analysis is done for the modeled system and time delays are known.

Comparing the models from volunteers and experts can be done in \figref{kapenc_compare_faulty}, where only the
expert got the right configuration, but all the resulting models are comparable, besides rotation.

@@

First let us recall various distance formulas. Comoving distances
involve integrals of the type
\begin{equation}
D = \int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
%\raise.5ex\hbox{.}
\end{equation}
Specifically, the comoving distance $D_S$ from observer to source,
$D_{LS}$ from lens to source, and $D_L$ from observer to lens are
given by the above integral, with limits as follows.
\begin{equation}
\begin{aligned}
&D_S    &                                &\int_0^{z_S} \\
\noalign{\medskip}
&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
\noalign{\medskip}
&D_L    &                                &\int_0^{z_L}
\end{aligned}
\end{equation}
Angular-diameter distances $d$ can be related to comoving distances
$D$ by scale factors, as follows.
\begin{equation}
\begin{aligned}
D_S &= (1+z_S) \, d_S \\
D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
D_L &= (1+z_L) \, d_L
\end{aligned}
\end{equation}
Consequently, angular-diameter distances are always smaller than
comoving distances.  Moreover, while comoving distances can be simply
added, angular distances cannot:
\begin{equation}
d_S \neq d_L + d_{LS}
\end{equation}

@@

>>>>

\begin{abstract}
In Spacewarps, volunteers are invited to search the CFHTLS sky survey for
gravitational lens candidates.  Here we report on a web interface    \todo[red]{or should that be application?} that allows experienced
volunteers to model lens candidates found within the classification phase of the project. A sample of 29 simulated lenses
were modeled multiple times by different users.  The quality of the models were then
examined in several ways.  Particular attention was paid to the following: (i)~the image parities and
time arrivals were identified correctly in most cases, but not
always; (ii)~the mean convergence (equivalent to the enclosed mass)
was quite well constrained in the image region; (iii)~the performance
of experienced volunteers was comparable to professionals.
\end{abstract}

@@

In response to the lessons learned from automated lens searches, the
{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org}
 has been launched.  {\em Spacewarps\/} is
part of the {\em Zooniverse\/} family of citizen-science
projects,\footnote{\tt http://www.zooniverse.org} where members of the
public are invited to analyze different kinds of scientific data which
are too difficult for robots and too large for specialists.  In {\em
 Spacewarps\/} itself, survey sky images are presented for volunteers to identify lens candidates.  The first tranche of survey data, introduced from May 2013, is an area of $\sim172$ square
degrees (or a quarter of a percent of the sky) from the Canada-France-Hawaii Telescope Legacy Survey.
 The survey is divided into patches of $440\times 440$ pixels, each of
which is seen by ten volunteers.  Simulated lenses are mixed in with
the data, both to help train volunteers on what to look for, and to
estimate completeness and purity.
These data had previously been analyzed by robots \cite{Gavazzi2012, More2012ApJ}.  
  A list of candidates is being
processed, but intermediate results show some very good candidates, as
well as re-discoveries of candidates previously found robotically.

@@

The sims were produced using {\tt gravlens}
\citep{2001astro.ph..2341K,2001astro.ph..2340K} using the CFHTLS survey data
and catalogues \cite{Coupon2009}. They were of three
kinds, as follows.

\begin{enumerate}
  \item Imitating lensed quasars: having a singular elliptical
    isothermal lens (SIE) plus constant external shear, and a circular
    Gaussian source.
  \item Emulating lensed galaxies: similar to the above, but with an
    elliptical de Vaucouleurs source.
  \item Resembling cluster lenses: having a source as above, but a
    more complicated lens, with one dominant elliptical SIE and
    one or more perturbing elliptical SIEs, plus a circular NFW
    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
    the underlying dark matter distribution.
\end{enumerate}

@@

To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate \ERg.
To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
These results can be seen in \figref{ER_per_sim}.
This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.

@@

\begin{figure}
  \centering
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007020_kappa_encl}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007024_kappa_encl}}
  \subfigure{\includegraphics[width=0.45\textwidth]
             {fig/007021_kappa_encl}} \\
  \caption{Enclosed mass \kenc for models of \asw{0h2m} with incorrect image identifications. \figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).  Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).}
\label{fig:kapenc_compare_faulty}
\end{figure}

@@

Comparing the models from volunteers and experts can be done in Figures \ref{fig:7022}, \ref{fig:7025} and \ref{fig:kapenc_compare_faulty}, where only the
expert got the right configuration (\figref{7022}) but enclosed masses are all fairly similar.  In cases with nearly symmetric image configurations, like this example, it is more difficult to identify the image parities correctly.  Incorrect image-parity identification changes the orientation of the mass distribution \kap[x,y],
but \kenc and thus the \ERf are not influenced that much.

@@

First let us recall various distance formulas. Comoving distances
involve integrals of the type
\begin{equation}
\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
%\raise.5ex\hbox{.}
\end{equation}
Specifically, the comoving distance $D_S$ from observer to source,
$D_{LS}$ from lens to source, and $D_L$ from observer to lens are
given by the above integral, with limits as follows.
\begin{equation}
\begin{aligned}
&D_S    &                                &\int_0^{z_S} \\
\noalign{\medskip}
&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
\noalign{\medskip}
&D_L    &                                &\int_0^{z_L}
\end{aligned}
\end{equation}
Angular-diameter distances $d$ can be related to comoving distances
$D$ by scale factors, as follows.
\begin{equation}
\begin{aligned}
D_S &= (1+z_S) \, d_S \\
D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
D_L &= (1+z_L) \, d_L
\end{aligned}
\end{equation}
Consequently, angular-diameter distances are always smaller than
comoving distances.  Moreover, while comoving distances can be simply
added, angular distances cannot:
\begin{equation}
d_S \neq d_L + d_{LS}
\end{equation}

@@

\section{Miscellaneous}

Maybe discard \figref{ER_all_models}, since the essential information is there in \figref{ER_per_sim}.
